---
title: "Back Calculated Length at Age"
author: "Anna Conklyn"
date: "3/8/2023"
output: html_document
---

# **Structure and Condition of an Established Round Goby Population in the Thousand Islands Region, St. Lawrence River**
## *Back Calculated Length at Age*

### Contents of Rmarkdown:
  > 1. Data setup, convert from one-fish-per-line to one-measurement-per-line
  > 2. Linear regression fitting of otolith radius vs. total length to aquire coefficients for back calculation
  > 3. Back-calculate total lengths at age
  > 4. Create summary tables and plots for predicted length-at-age
  
```{r setup, include=FALSE}
##Install packages

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(FSA)
library(nlstools)
library(magrittr)
library(dplyr)
library(plotrix)
library(nnet)
library(nls2)
library(ggplot2)
library(tidyr)
library(broom)
library(stargazer)
library(plyr)
library(gt)
library(gtsummary)
library(stringr)
library(lattice)
library(car)
library(multcomp)
library(readxl)
```


```{r, include=FALSE}
##Load data


getwd()
setwd("G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE")
RG <- read_excel("Pop_Strucutre_Dissected.xlsx")

RG<- filter(RG, Age != "NA")
```



#### 1. Data setup, convert from one-fish-per-line to one-measurement-per-line

```{r, include=FALSE}
RG$yearclass<-RG$Year-RG$Age

Male<- filter(RG, Sex== "M")
Female<- filter(RG, Sex=="F")
```


```{r, include=FALSE}
RG_long_M<- Male %>% pivot_longer(cols = c( 'RadAnnu0', 'RadAnnu1', 'RadAnnu2', 'RadAnnu3', 'RadAnnu4', 'RadAnnu5'), names_to = "AgeRad", values_to = "Rad")
RG_long_M$AgeRad<- str_remove_all(RG_long_M$AgeRad, "[RadAnnu]")
RG_long_M$AgeRad<- as.numeric(RG_long_M$AgeRad)

RG_long_F<- Female %>% pivot_longer(cols = c( 'RadAnnu0', 'RadAnnu1', 'RadAnnu2', 'RadAnnu3', 'RadAnnu4', 'RadAnnu5'), names_to = "AgeRad", values_to = "Rad")
RG_long_F$AgeRad<-str_remove_all(RG_long_F$AgeRad, "[RadAnnu]")
RG_long_F$AgeRad<- as.numeric(RG_long_F$AgeRad)
```

### 2. Linear regression fitting of otolith radius vs. total length to aquire coefficients for back calculation

#### Male
```{r, echo=FALSE}
BodyOR_M<- lm(log(Total.Length) ~ log(OR), data=Male) 
(BodyOR_M_Summary<- coef(BodyOR_M))
a_M<- coef(BodyOR_M)[2]
```

#### Female
```{r, echo=FALSE}
BodyOR_F<- lm(log(Total.Length) ~ log(OR), data=Female) 
(BodyOR_F_Summary<- coef(BodyOR_F))
a_F<- coef(BodyOR_M)[2]
```

### 3. Back-calculate total lengths at age

$$ TL_{n} = TL * (OR_{n}/OR)^a $$
TL_{n} is the total length at annulus n; TL, observed total length; OR_{n} otolith radius at annulus n; OR, otolith radius at capture; a, the slope of linear function between logarithm of total length and logarithm of total radius 

#### Male
```{r, echo=FALSE}
RG_long_M$Pred_len<- with(RG_long_M, ((Rad/OR)*Total.Length)^a_M)
headtail(RG_long_M)
```

#### Female
```{r, echo=FALSE}
RG_long_F$Pred_len<- with(RG_long_F, ((Rad/OR)*Total.Length)^a_F)
headtail(RG_long_F)
```

### Create summary tables and plots for predicted length-at-age

#### Male
```{r, echo=FALSE}
M_BC_Sum<- Summarize(Pred_len~AgeRad,data=RG_long_M,digits=2)
M_BC_Sum$Sex<- "Male"
RG_long_M$Pred_len<- as.numeric(RG_long_M$Pred_len)
RG_long_M$AgeRad<- as.numeric(RG_long_M$AgeRad)

(Male_BC_table =  gt(M_BC_Sum[1:4], rowname_col = "AgeRad") %>% cols_hide(columns="n") %>% tab_header(title="Male") %>% cols_label(
    nvalid = "n",
    mean = "mean") %>% cols_align(align = "left") %>%
  tab_stubhead(label = "Age") %>% tab_options(table.width = pct(50))%>% cols_width(everything() ~ px(100)))

(M_BC_jitter<- ggplot(RG_long_M, aes(AgeRad, Pred_len)) + geom_jitter(aes(colour=Site)) + facet_wrap(vars(Age)) + ylim (0,300) + xlim(0,5) + ylab ("Back-Calculated Length (mm)") + xlab ("Age (Jittered)") + labs(title = "Male") + theme_classic () + scale_color_manual(values = c("Rose" = "black", "Cobb" = "grey")) +  theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 15), legend.position = "bottom"))

M_BC_jitter %>% ggsave(filename="Male BC_jitter.png", path="G:/POP_STRUCTURE_PUB/Figures", dpi="retina", width = 7.48031, height = 5)

```


```{r, echo=FALSE}
###Check ANCOVA assumptions## 
summary(aov(AgeRad~Site, data=RG_long_M)) ## covariate and treatment are independent##
RG_long_M$AgeRad <- as.factor(RG_long_M$AgeRad) 
leveneTest(Pred_len ~ AgeRad, data = RG_long_M) ##test for homogeneity of variance among sites

 
MaleBC <- aov(Pred_len ~ AgeRad + Site, data = RG_long_M)
(shapiro.test(resid(MaleBC)))
MaleBC_table<-(Anova(MaleBC, type=3))
rownames(MaleBC_table)[rownames(MaleBC_table) == "AgeRad"] <- "Age"
MaleBC_table$rownames<- rownames(MaleBC_table)

(MaleBC_table_print <- MaleBC_table %>% gt(rowname_col = "rownames") %>% fmt_number(columns = c("Sum Sq" , "F value"),  decimals = 2)  %>% tab_header( title = md("**Male**")))

MaleBC_table_print %>% gtsave(filename="Male BC_ANCOVA.png", path="G:/POP_STRUCTURE_PUB/Figures")
```


#### Female
```{r, echo=FALSE}
F_BC_Sum<- Summarize(Pred_len~AgeRad,data=RG_long_F,digits=2)
F_BC_Sum$Sex<- "Female"
RG_long_F$Pred_len<- as.numeric(RG_long_F$Pred_len)
RG_long_F$AgeRad<- as.numeric(RG_long_F$AgeRad)

(Female_BC_table = gt(F_BC_Sum[1:5], rowname_col = "AgeRad") %>% cols_hide(columns="n") %>% tab_header(title="Female") %>% cols_label(
    nvalid = "n",
    mean = "mean") %>% cols_align(align = "left") %>%
  tab_stubhead(label = "Age") %>% tab_options(table.width = pct(50))%>% cols_width(everything() ~ px(100)))

(F_BC_jitter<- ggplot(RG_long_F, aes(AgeRad, Pred_len)) + geom_jitter(aes(colour=Site)) + facet_wrap(vars(Age)) + ylim (0,300) + xlim(0,5) + ylab ("Back-Calculated Length (mm)") + xlab ("Age (Jittered)") + labs(title = "Female") + theme_classic () + scale_color_manual(values = c("Rose" = "black", "Cobb" = "grey")) +  theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 15), legend.position = "bottom"))

 F_BC_jitter %>% ggsave(filename="Female BC_jitter.png", path="G:/POP_STRUCTURE_PUB/Figures")

```

```{r}
(BCplot <- ggarrange(M_BC_jitter+ rremove ("ylab"), F_BC_jitter+ rremove ("ylab"),
                    nrow = 2, common.legend = TRUE, legend="bottom"))

(BCplot<- annotate_figure(BCplot,left=text_grob("Back-Calculated Length (mm)", rot = 90, size = 15)))

ggsave("BCplot.jpeg", device="jpeg", plot=BCplot, path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures", dpi="retina", width = 7.48031, height = 9)
```




```{r, echo=FALSE}

###Check ANCOVA assumptions

summary(aov(AgeRad~Site, data=RG_long_F)) ## covariate and treatment are independent

RG_long_F$AgeRad <- as.factor(RG_long_F$AgeRad) 
leveneTest(Pred_len ~ AgeRad, data = RG_long_F) ##test for homogeneity of variance among sites

 
FemaleBC <- aov(Pred_len ~ AgeRad + Site, data = RG_long_F)
(shapiro.test(resid(FemaleBC)))
FemaleBC_table<-(Anova(FemaleBC, type=3))
rownames(FemaleBC_table)[rownames(FemaleBC_table) == "AgeRad"] <- "Age"
FemaleBC_table$rownames<- rownames(FemaleBC_table)

(FemaleBC_table_print <- gt(FemaleBC_table, rowname_col = "rownames") %>% fmt_number(columns = c("Sum Sq" , "F value"),  decimals = 2) %>% tab_header( title = md("**Female**")))

FemaleBC_table_print %>% gtsave(filename="Female BC_ANCOVA.png", path="G:/POP_STRUCTURE_PUB/Figures")
 

MaleBC_table$Sex<-"Male"
FemaleBC_table$Sex<-"Female"

FullBC_ANCOVA_table<- rbind(MaleBC_table, FemaleBC_table)

FullBC_ANCOVA<- gt(FullBC_ANCOVA_table, rowname_col = "rownames", groupname_col = "Sex") %>% fmt_number(columns = c("Sum Sq" , "F value", "Pr(>F)"),  decimals = 2) %>% 
  tab_style(
    style = cell_text( weight = "bold", align = "left"),
    locations = cells_group()) %>% 
  tab_style(
    style = cell_text( weight = "bold", align = "left"),
    locations = cells_column_labels()
  )
  
FullBC_ANCOVA %>% gtsave(filename = "BCANCOVA table.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures")

```



```{r}

FullBC<- rbind(M_BC_Sum, F_BC_Sum)

(FullBC_Table= gt(FullBC[1:5], rowname_col = "AgeRad") %>% cols_hide(columns="n") %>% tab_header(title= md("**Mean back-calculated length-at-age**")) %>% cols_label(
    nvalid = md("**n**"),
    mean = md("**Mean**"), sd = md("**SD**")) %>% cols_align(align = "left") %>%
  tab_stubhead(label = md("**Age**")) %>% tab_options(table.width = pct(50)) %>% cols_width(everything() ~ px(100)) %>%
  tab_row_group(
    group = md("**Female**"),
    rows = 7:12
  )  %>% tab_row_group(
    group = md("**Male**"),
    rows = 1:6
  ) )

 
FullBC_Table %>% gtsave(filename="BC_Table.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures")
```



