---
title: "Age Summary and Body OR"
author: "Anna Conklyn"
date: "9/14/2021"
output: html_document
---

# **Structure and Condition of an Established Round Goby Population in the Thousand Islands Region, St. Lawrence River**
## *Age summary and Body-Otolith Radius relationship*

### Contents of Rmarkdown:
  > 1. Summary table and histogram of round goby ages for both sexes and sites
  > 2. Body otolith regression and statistical analyses
  > 3. 



```{r setup, include=FALSE}
##Install packages

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(FSA)
library(readxl)
library(nlstools)
library(magrittr)
library(dplyr)
library(plotrix)
library(nnet)
library(nls2)
library(ggplot2)
library(tidyr)
library(broom)
library(stargazer)
library(plyr)
library(gt)
library(gtsummary)
library(ggpubr)
library(gtExtras)
library(sjPlot)
```


```{r, include=FALSE}
##Load data

getwd()
RG <- read_excel("C:/Users/ah2355/Documents/GitHub/RG_POP_STRUCTURE/Data/Pop_Strucutre_Dissected.xlsx")
```



```{r, include=FALSE}
#5mm & 10mm length categories; remove unaged

RG %<>% mutate(lcat5=lencat(Total.Length, w=5)) %<>% mutate(lcat10=lencat(Total.Length, w=10)) %<>% filter(!is.na(RG$Age))  %<>% mutate(logTL=log10(Total.Length), logOR=log10(OR), Site=factor(Site)) 

RG<- filter(RG, Sex!="I")

headtail(RG)

Male<- filter(RG, Sex== "M")
Female<- filter(RG, Sex=="F")

Rose<- filter(RG, Site=="Rose")
Cobb<- filter(RG, Site=="Cobb")
```


## 1. Summary histogram of round goby ages for both sexes and sites

### Hisogram of ages

```{r, echo=FALSE}

RG$Sex <- factor(RG$Sex, levels = c("M", "F", "I"),  labels = c("Male", "Female", "Immature"))

(Age_plot<-ggplot(RG, aes(x=Age, y=..count.., fill=Site)) + geom_bar(binwidth=1, color="black", position=position_dodge(0.9, preserve = "single")) + facet_grid(cols = vars(RG$Sex)) + labs(x = "Age",  y = "Count") + scale_fill_manual(values=c("grey", "black")) + theme_classic () + xlim(0,5)+ theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 12), legend.position = "bottom"))

(MAge_plot<-ggplot(Male, aes(x=Age, y=..count.., fill=Site)) + geom_bar(color="black", position=position_dodge(0.9, preserve = "single")) + labs(x = "Age",  y = "Count", title = "Male") + ylim(0,80) +  scale_fill_manual(values=c("grey", "black")) + theme_classic () + xlim(0,5) + theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 12), legend.position = "bottom"))

(FAge_plot<-ggplot(Female, aes(x=Age, y=..count.., fill=Site)) + geom_bar(color="black", position=position_dodge(0.9, preserve = "single")) + labs(x = "Age",  y = "Count", title = "Female") + ylim(0,80) +  scale_fill_manual(values=c("grey", "black"))+ theme_classic () + xlim(0,5) + theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 12), legend.position = "bottom"))

Ageplot <- ggarrange(MAge_plot + rremove("ylab"), FAge_plot+ rremove("ylab"),
                    nrow = 2, common.legend = TRUE, legend="bottom")

(Ageplot<-annotate_figure(Ageplot,left=text_grob("Count", rot = 90, size = 12)))

(Age_Histogram<-ggplot(RG, aes(x=Age, y=..count.., fill=Sex)) + geom_bar(color="black", position=position_dodge(1, preserve = "single")) + labs(x = "Age",  y = "Count") +  scale_fill_manual(values=c("grey", "black"))+ theme_classic () + xlim(1,5.5) + theme(text = element_text(size = 12), legend.position = "bottom"))


ggsave("Age_Histogram.tiff", device="tiff", plot=Age_Histogram, path="C:/Users/ah2355/Documents/GitHub/RG_POP_STRUCTURE/Figures/TIFFs", dpi="retina", width = 3.3, height = 3)
```


### Summary tables of round goby ages by length category for each sex
```{r echo=FALSE}

library(kableExtra)
MaleAge<- xtabs(~  Age + lcat10, data=Male)
MaleAge<- as.data.frame.matrix(MaleAge)
MaleAge<-t(MaleAge)
MaleAge<- as.data.frame(MaleAge)
MaleAge$Lcat<-row.names(MaleAge)
(MaleAge_Table<- gt(MaleAge, rowname_col = "Lcat") %>%
    tab_stubhead(label = "Length Interval (cm)"))
colnames(MaleAge)<- c("Male.1",    "Male.2" ,   "Male.3" ,   "Male.4" ,   "Male.5" ,   "Lcat")
MaleAge$Lcat<- as.numeric(MaleAge$Lcat)
MaleAge$Lcat<-sort(MaleAge$Lcat, decreasing = FALSE)



FemaleAge<- xtabs(~  Age + lcat10, data=Female)
FemaleAge<- as.data.frame.matrix(FemaleAge)
FemaleAge<-t(FemaleAge)
FemaleAge<- as.data.frame(FemaleAge)
FemaleAge$Lcat<-row.names(FemaleAge)
(FemaleAge_Table<- gt(FemaleAge, rowname_col = "Lcat") %>%
    tab_stubhead(label = "Length Interval (cm)"))
colnames(FemaleAge)<- c("Female.1",    "Female.2" ,   "Female.3" ,   "Female.4" ,   "Female.5" ,   "Lcat")
FemaleAge$Lcat<- as.numeric(FemaleAge$Lcat)
FemaleAge$Lcat<-sort(FemaleAge$Lcat, decreasing = FALSE)

Age_Length<- merge(MaleAge, FemaleAge, by.x="Lcat", all = TRUE)


Age_Length$Lcat<- as.numeric(Age_Length$Lcat)
Age_Length$Lcat<-sort(Age_Length$Lcat, decreasing = FALSE)


(Age_Length_table<- gt (Age_Length, rowname_col = "Lcat", groupname_col = "Sex") %>%
   tab_stubhead(label = "Length Interval (mm)") %>%
   tab_spanner_delim(delim = ".") %>% 
   tab_options(column_labels.font.weight = "bold", row_group.font.weight = "bold", table.align = "left") %>% 
  sub_missing(
  columns = everything(),
  rows = everything(),
  missing_text = ("0"))  %>%  
    tab_header(
    title = md ("**Age group (years)**" )) %>%
  summary_rows(columns = everything(), fns = list(Total = ~sum(., na.rm = TRUE)), formatter = fmt_number, decimals = 0) %>% 
  cols_width(vars(Lcat) ~ px(100),
    everything() ~ px(60)
  ) %>% cols_hide(columns = "Lcat") )

Age_Length_table %>% gtsave(filename="Age_Length_table.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures")


```


## 2. Body otolith regression and statistical analyses

### Scatterplot of otolith radius vs. total length
```{r, results='asis', echo=FALSE}
RG$Sex<- factor(RG$Sex, levels = c("M", "F"), labels= c("Male", "Female"))

ggplot(RG, aes(x=logOR, y=logTL)) + geom_point(aes(color=Site)) + facet_grid (rows=vars(Sex)) + theme_classic() + scale_color_manual(values=c("gray", "black")) + labs(x= "Log(Otolith Radius) (mm)", y="Log (Total Length) (mm)")+ theme(legend.position="bottom")+ geom_smooth(method = "lm", aes(group = Site, color=Site, linetype=Site))


(MaleOR_plot<-ggplot(Male, aes(x=logOR, y=logTL)) + geom_point(aes(color=Site)) + theme_classic() + scale_color_manual(values=c("gray", "black")) + labs(x= "Log(Otolith Radius) (mm)", y="Log (Total Length) (mm)", title = "Male")+ theme(legend.position="bottom")+ geom_smooth(method = "lm", aes(group = Site, color=Site, linetype=Site)) + theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 12), legend.position = "bottom") + ylim(1,3)+ xlim(0,0.6))


(FemaleOR_plot<-ggplot(Female, aes(x=logOR, y=logTL)) + geom_point(aes(color=Site)) + theme_classic() + scale_color_manual(values=c("gray", "black")) + labs(x= "Log(Otolith Radius) (mm)", y="Log (Total Length) (mm)", title = "Female")+ theme(legend.position="bottom")+ geom_smooth(method = "lm", aes(group = Site, color=Site, linetype=Site)) + theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 12), legend.position = "bottom") + ylim(1,3) + xlim(0,0.6))

(ORplot <- ggarrange(MaleOR_plot + rremove ("ylab"), FemaleOR_plot + rremove ("ylab"),
                    nrow = 2, common.legend = TRUE, legend="bottom"))

(ORplot<-annotate_figure(ORplot,left=text_grob("Log (Total Length) (mm)", rot = 90, size = 12)))

ggsave("OR_TL.tiff", device="tiff", plot=ORplot, path="C:/Users/ah2355/Documents/GitHub/RG_POP_STRUCTURE/Figures/TIFFs", dpi="retina", width = 6.85, height = 5)

```


### Linear regression fitting of otolith radius vs. total length, Summary tables and F tests for equality of slope between sexes and sites

#### Differences between sites for each sex
```{r, echo=FALSE, results='asis'}

M_BodyOR<- lm(logTL ~ logOR * Site, data=Male) 
(MSummary<- summary(M_BodyOR))

F_BodyOR<- lm(logTL ~ logOR * Site, data=Female) 
(FSummary<- summary(F_BodyOR))

(MOR_Table<- tbl_regression(M_BodyOR) %>%
  add_glance_table(
    include = c(nobs, r.squared)) %>%  
  add_significance_stars(hide_p = FALSE,
                           pattern = "{p.value}{stars}"))

FOR_Table<- tbl_regression(F_BodyOR) %>%
  add_glance_table(
    include = c(nobs, r.squared))%>%  
  add_significance_stars(hide_p = FALSE,
                           pattern = "{p.value}{stars}") 

Sex_OR_Table<- tbl_merge(
    tbls = list(MOR_Table, FOR_Table),
    tab_spanner = c("**Male**", "**Female**")
  ) %>% as_gt()
 
 Sex_OR_Table %>% gtsave(filename="Sex_OR_Table.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures")

```

#### Differences between sexes for each site
```{r, echo=FALSE, results='asis'}
Rose_BodyOR<- lm(logTL ~ logOR * Sex, data=Rose) 
RoseORSummary<- summary(Rose_BodyOR)

Cobb_BodyOR<- lm(logTL ~ logOR * Sex, data=Cobb)
CobbORSummary<- summary(Cobb_BodyOR)

(RoseOR_Table<- tbl_regression(Rose_BodyOR) %>%
  add_glance_table(
    include = c(nobs, r.squared)) %>%  
  add_significance_stars(hide_p = FALSE,
                           pattern = "{p.value}{stars}"))

CobbOR_Table<- tbl_regression(Cobb_BodyOR) %>%
  add_glance_table(
    include = c(nobs, r.squared))%>%  
  add_significance_stars(hide_p = FALSE,
                           pattern = "{p.value}{stars}") 

Site_OR_Table<- tbl_merge(
    tbls = list(RoseOR_Table, CobbOR_Table),
    tab_spanner = c("**Rose**", "**Cobb**")
  ) %>% as_gt()
 
 Site_OR_Table %>% gtsave(filename="Site_OR_Table.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures")
```

# Male and Female regression parameters

```{r}
M_BodyOR<- lm(logTL ~ logOR, data=Male) 
(MSummary<- summary(M_BodyOR))

F_BodyOR<- lm(logTL ~ logOR, data=Female) 
(FSummary<- summary(F_BodyOR))
```

