---
title: "Compiled_Condition"
author: "Anna Conklyn"
date: "2/28/2021"
output: html_document
---

# **Structure and Condition of an Established Round Goby Population in the Thousand Islands Region, St. Lawrence River**
## *Condition indices*

### Contents of Rmarkdown:
  > 1. Calculate Condition Indices
  > 2. Monthly condition index analysis: (HSI, GSI, K)
  >       a.Normality tests (Shapiro Wilk) & Kruskal Wallis post hoc multiple comparisons
  >       b.Summary statistics and monthly plot creation (by sex)
          c. M-F plot bind
  
  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




```{r, include=FALSE}

library(readxl)
library(betareg)
library(plyr)
library(lmtest)
library(glmmTMB)
library(boot)
library(emmeans)
library(ggplot2)
library(mvtnorm)
library(nlme)
library(tidyr)
library(dplyr)
library(magrittr)
library(ggpubr)
library(lme4)
library(lmerTest)
library(FSA)
library("viridis")   
library(rstatix)
library(ggprism)
library(ggtext)
library(Rmisc)
```


```{r, include = FALSE}
AllRG <- read_excel("C:/Users/ah2355/Documents/GitHub/RG_POP_STRUCTURE/Data/Pop_Strucutre_Dissected.xlsx")
```


## 1. Calculate condition indices

#### $$ HSI = ( Liver Mass / Total Mass ) * 100 $$
#### $$ GSI = ( Gonad Mass / Total Mass ) * 100 $$
#### $$ K = ( Total Mass / Total Length^3 ) * 10^5 $$

```{r, echo=FALSE}
#Calculate indices, filter sexes 

AllRG$Month <- factor(AllRG$Month , levels=c("May", "June", "July", "August", "September", "October"))
AllRG<- AllRG %>% filter(Month != "June") %>% filter(Month != "September") %>% filter(Sex != "I")

AllRG$Liver.Mass<- as.numeric(AllRG$Liver.Mass)

AllRG %<>% mutate(HSI=(Liver.Mass/Total.Mass), GSI=(Gonad.Mass/Total.Mass), K=(Total.Mass/Total.Length^3)) %<>% mutate (Sex=factor(Sex), Site=factor(Site))


RG.M<-AllRG %>% filter(Sex=="M")
RG.F<-AllRG %>% filter(Sex=="F")

```


## 2. Monthly condition index analysis: GSI

### Male GSI

#### a. Normality tests (Shapiro Wilk) & Kruskal Wallis and dunn test post hoc multiple comparison
```{r, echo=FALSE}
shapiro.test(RG.M$GSI)
kruskal.test(GSI ~ Month, data = RG.M)
(MGSI_Comp<- dunn_test(GSI ~ Month, data = RG.M, p.adjust.method = "holm"))
```

#### b. Summary statistics and monthly plot creation

```{r, echo=FALSE}
(M_GSI_Summary<-(summarySE(RG.M, measurevar="GSI", groupvars=c("Site", "Month"), na.rm = TRUE)))

(GSI_M_line<- ggplot(M_GSI_Summary, aes(y=GSI*100, x=Month, colour=Site)) + ylim(0, 10) + geom_errorbar(aes(ymin=(GSI-se)*100, ymax=(GSI+se)*100), width=.1)+ labs(title = "Male") + geom_point() + geom_line(aes(group=Site))+ theme_classic () + scale_colour_manual(values=c( "grey","black"))  +geom_hline(yintercept=1, linetype="dashed", size=1) + theme(text = element_text(size = 12), legend.position = "bottom") + ylab("Gonadosomatic Index \n (Gonad Weight/Total Weight * 100)"))

ggsave("GSI_Lineplot_male.jpeg", device="jpeg", plot=GSI_M_line, path="G:/POP_STRUCTURE_PUB/Figures", dpi="retina", width = 6.5, height = 5)

```


```{r, echo=FALSE}
M_GSI<- ggplot(RG.M, aes(x=Month, y=GSI*100, fill=Site)) + geom_boxplot(position=position_dodge(1)) + labs(y=NULL, x = NULL, title = "Male", fill = "Site") + ylim(0, 20) + scale_fill_manual(values=c( "grey" ,"black" )) +geom_hline(yintercept=1, linetype="dashed", size=1) + theme (axis.text.x = element_text( size = 12), axis.text.y = element_text( size = 12), axis.title.x = element_text(size = 15), axis.title.y = element_text( size = 13), legend.text = element_text(size=13), legend.title = element_text(size=15)) +theme_classic ()
M_GSI
```



### Female GSI

#### a. Normality tests (Shapiro Wilk) & Kruskal Wallis and dunn test post hoc multiple comparison
```{r, echo=FALSE}
shapiro.test(RG.F$GSI)
kruskal.test(GSI ~ Month, data = RG.F)
(FGSI_Comp<- dunn_test(GSI ~ Month, data = RG.F, p.adjust.method = "holm"))
```


#### b. Summary statistics and monthly plot creation

```{r, echo=FALSE}
(F_GSI_Summary<-(summarySE(RG.F, measurevar="GSI", groupvars=c("Site", "Month"), na.rm = TRUE)))

(GSI_F_line<- ggplot(F_GSI_Summary, aes(y=GSI*100, x=Month, colour=Site)) + ylim(0, 10) + geom_errorbar(aes(ymin=(GSI-se)*100, ymax=(GSI+se)*100), width=.1) + labs(title = "Female") + geom_point() + geom_line(aes(group=Site))+ theme_classic () + scale_colour_manual(values=c( "grey","black")) + geom_hline(yintercept=5, linetype="dashed", size=1)  + theme(text = element_text(size = 12), legend.position = "bottom") + ylab("Gonadosomatic Index \n (Gonad Weight/Total Weight * 100)"))

ggsave("GSI_Lineplot_female.tiff", device="tiff", plot=GSI_F_line, path="C:/Users/ah2355/Documents/GitHub/RG_POP_STRUCTURE/Figures/TIFFs", dpi="retina", width = 6.5, height = 5)

```


```{r, echo=FALSE}
(F_GSI<- ggplot(RG.F, aes(x=Month, y=GSI*100, fill=Site)) + geom_boxplot(position=position_dodge(1)) + labs( y=NULL, x = NULL, title = "Female", fill = "Site") + ylim(0, 20) + scale_fill_manual(values=c("black", "Grey"))  +  geom_hline(yintercept=5, linetype="dashed", size=1) + theme (axis.text.x = element_text( size = 12), axis.text.y = element_text( size = 12), axis.title.x = element_text( size = 12), axis.title.y = element_text( size = 13), legend.text = element_text(size=13), legend.title = element_text(size=15)) +theme_classic ())

```



## M-F plot bind

```{r, echo=FALSE}
GSI_LINE<- ggarrange(GSI_M_line + rremove ("ylab"), GSI_F_line + rremove ("ylab") , ncol = 1, nrow = 2, common.legend = TRUE, legend = "bottom")

(GSI_LINE<- annotate_figure(GSI_LINE, left= text_grob("Gonadosomatic Index \n (Gonad Weight/Total Weight * 100)", rot = 90, size = 12)))
```


```{r}

(GSI_Summary<-(summarySE(AllRG, measurevar="GSI", groupvars=c("Sex", "Month"), na.rm = TRUE)))

(GSI_LINE<- ggplot(GSI_Summary, aes(y=GSI*100, x=Month, colour=Sex)) + ylim(0, 10) + geom_errorbar(aes(ymin=(GSI-se)*100, ymax=(GSI+se)*100), width=.1) + geom_point() + geom_line(aes(group=Sex))+ theme_classic () + scale_colour_manual(values=c("black", "grey")) + geom_hline(yintercept=5, linetype="dashed", size=1, color="black")+ geom_hline(yintercept=1, linetype="dashed", size=1, color="grey")  + theme(text = element_text(size = 12), legend.position = "bottom") + ylab("Gonadosomatic Index \n (Gonad Weight/Total Weight \n * 100)") + theme(legend.title=element_blank(), legend.margin=margin(c(0,1,1,1))))

ggsave("GSI_LINE.tiff", device="tiff", plot=GSI_LINE, path="C:/Users/ah2355/Documents/GitHub/RG_POP_STRUCTURE/Figures/TIFFs", dpi="retina", width = 3.3, height = 3)
```

```{r}
(M_GSI_Scatter<- ggplot(data = RG.M) + geom_point(aes(x=Total.Length, y=GSI*100, color=Month, shape = Site)) + labs(y="GSI (Gonad Weight/Total Weight * 100)", x="Total Length (mm)", title = "Male") + ylim(0, 20) + geom_hline(yintercept=1, linetype="dashed", size=1) + scale_colour_brewer(palette ="RdBu" , name = "Month", labels = c("May", "July", "August", "October"), direction = -1 )+theme_classic () + theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 12), legend.position = "bottom"))

(F_GSI_Scatter<- ggplot(data = RG.F) + geom_point(aes(x=Total.Length, y=GSI*100, color=Month, shape = Site)) + labs(y=NULL, x="Total Length (mm)", title = "Female") + ylim(0, 20) +  geom_hline(yintercept=5, linetype="dashed", size=1) +  scale_colour_brewer(palette ="RdBu" , name = "Month", labels = c("May", "July", "August", "October"), direction = -1 )+theme_classic ()+ theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 12), legend.position = "bottom"))

(GSI_Scatter<- ggarrange(M_GSI_Scatter, F_GSI_Scatter,
                    ncol = 2, common.legend = TRUE, legend="bottom"))

 
 ggsave("JOINED_GSI.jpeg", device="jpeg", plot=GSI_Scatter, path="G:/POP_STRUCTURE_PUB/Figures", dpi="retina", width = 7.48031, height = 5)
```


# 2. Monthly condition index analysis: HSI

### Male HSI

#### a. Normality tests (Shapiro Wilk) & Kruskal Wallis and dunn test post hoc multiple comparison

```{r, echo=FALSE}

shapiro.test(RG.M$HSI)
kruskal.test(HSI ~ Month, data = RG.M)
(MHSI_Comp<- dunn_test(HSI ~ Month, data = RG.M, p.adjust.method = "holm"))

```

#### b. Summary statistics and monthly plot creation

```{r, echo=FALSE}
(M_HSI_Summary<-(summarySE(RG.M, measurevar="HSI", groupvars=c("Site", "Month"))))
M_HSI_Summary<- M_HSI_Summary %>% filter(Month != "June") %>% filter(Month != "September")

(HSI_M_line<- ggplot(M_HSI_Summary, aes(y=HSI*100, x=Month, colour=Site))  + ylab("Hepatosomatic Index \n (Liver Weight/Total Weight * 100)") + labs(title = "Male") + geom_errorbar(aes(ymin=(HSI-se)*100, ymax=(HSI+se)*100), width=.1) + geom_point() + geom_line(aes(group=Site))+ theme_classic () + scale_colour_manual(values=c( "grey","black")) + ylim(0, 6) + theme(plot.margin = margin(t=0.5, l=0.5, r=0.5, b=0.5, unit = "cm"), text = element_text(size = 15), legend.position = "bottom"))


```


```{r, echo=FALSE}
(HSI_M_box<- ggplot(RG.M, aes(x=Month, y=HSI*100, fill=Site)) + geom_boxplot(position=position_dodge(1)) + labs(y=NULL, x = NULL, title = "Male") + ylim(0, 8) + scale_fill_manual(values=c("black", "grey")) + labs(fill = "Sex") + theme(legend.position = c(0.8, 0.8), axis.text.x = element_text( size = 12), axis.text.y = element_text( size = 12), axis.title.x = element_text( size = 15), axis.title.y = element_text( size = 13), legend.text = element_text(size=13), legend.title = element_text(size=15))+theme_classic ())

```


### Female HSI

#### a. Normality tests (Shapiro Wilk) & Kruskal Wallis and dunn test post hoc multiple comparison
```{r, echo=FALSE}
shapiro.test(RG.F$HSI)
kruskal.test(HSI ~ Month, data = RG.F)
(FHSI_Comp<- dunn_test(HSI ~ Month, data = RG.F, p.adjust.method = "holm"))
```

#### b. Summary statistics and monthly plot creation

```{r, echo=FALSE}
(F_HSI_Summary<-(summarySE(RG.F, measurevar="HSI", groupvars=c("Site", "Month"))))

(HSI_F_line<- ggplot(F_HSI_Summary, aes(y=HSI*100, x=Month, colour=Site)) + ylim(0, 6) + ylab("Hepatosomatic Index \n (Liver Weight/Total Weight * 100)") + labs(title = "Female") + geom_errorbar(aes(ymin=(HSI-se)*100, ymax=(HSI+se)*100), width=.1) + geom_point() + geom_line(aes(group=Site))+ theme_classic () + scale_colour_manual(values=c( "grey","black")) + theme(plot.margin = margin(t=0.5, l=0.5, r=0.5, b=0.5, unit = "cm"), text = element_text(size = 15), legend.position = "bottom"))

```

```{r, echo=FALSE}
(HSI_F_box<- ggplot(RG.F, aes(x=Month, y=HSI*100, fill=Site)) + geom_boxplot(position=position_dodge(1)) + labs(y=NULL, x = NULL, title = "Female") + ylim(0, 8) + scale_fill_manual(values=c("black", "grey")) + labs(fill = "Sex") + theme(legend.position = c(0.8, 0.8), axis.text.x = element_text( size = 12), axis.text.y = element_text( size = 12), axis.title.x = element_text( size = 15), axis.title.y = element_text( size = 13), legend.text = element_text(size=13), legend.title = element_text(size=15)) +theme_classic ())

```


## M-F plot bind

```{r, echo=FALSE}
HSI_LINE<- ggarrange(HSI_M_line + rremove ("ylab"), HSI_F_line + rremove ("ylab"), ncol = 1, nrow = 2, common.legend = TRUE, hjust = c(-0.8, -0.5), legend = "bottom")

(HSI_LINE<- annotate_figure(HSI_LINE, left= text_grob("Hepatosomatic Index \n (Liver Weight/Total Weight * 100)", rot = 90, size = 15)))

ggsave("HSI_LINE.jpeg", device="jpeg", plot=HSI_LINE, path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures", dpi="retina", width = 7.48031, height = 9)
```



# 2. Monthly condition index analysis: K


```{r}
# Add not-dissected 
setwd("G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE")
RG_Dissected <- read_excel("Pop_Strucutre_Dissected.xlsx")

RG_Dissected<- RG_Dissected %>% dplyr:: select("Date", "Total.Mass", "Total.Length", "Sex", "Site", "Month", "Year")


RG_NotDissected <- read_excel("Pop_Strucutre_NotDissected.xlsx")
RG_NotDissected<- RG_NotDissected %>% dplyr:: select("Date", "Total.Mass", "Total.Length", "Sex", "Site", "Month", "Year")


AllRG<- rbind(RG_Dissected, RG_NotDissected)

as.factor(AllRG$Month)

AllRG$Month <- factor(AllRG$Month , levels=c("May", "June", "July", "August", "September", "October"))
AllRG<- AllRG %>% filter(Month != "June") %>% filter(Month != "September")
AllRG<- AllRG %>% filter(Total.Mass != "NA")


AllRG %<>% mutate(K=(Total.Mass/Total.Length^3)) %<>% mutate (Sex=factor(Sex), Site=factor(Site))


RG.M<-AllRG %>% filter(Sex=="M")
RG.F<-AllRG %>% filter(Sex=="F")
```

## Male K

#### a. Normality tests (Shapiro Wilk) & Kruskal Wallis and dunn test post hoc multiple comparison

```{r, echo=FALSE}
shapiro.test(RG.M$K)
kruskal.test(K ~ Month, data = RG.M)
(K_M_Comp<- dunn_test(K ~ Month, data = RG.M, p.adjust.method = "holm"))
```

#### b. Summary statistics and monthly plot creation

```{r, echo=FALSE}
(M_K_Summary<-summarySE(RG.M, measurevar="K", groupvars=c("Site", "Month"), na.rm = TRUE))

(K_M_line<- ggplot(M_K_Summary, aes(y=K*100000, x=Month, colour=Site)) + ylim(1, 2) + geom_errorbar(aes(ymin=(K-se)*100000, ymax=(K+se)*100000), width=.1) + geom_point() + labs(title = "Male") + geom_line(aes(group=Site))+ theme_classic () + scale_colour_manual(values=c("grey", "black")) + theme(plot.margin = margin(t=0.5, l=0.5, r=0.5, b=0.5, unit = "cm"), text = element_text(size = 15), legend.position = "bottom") + labs(y=bquote("Fulton's Condition \n (Total Weight / Total" *Length^3~")*" * 10^5~")")))

ggsave("K_Lineplot_male.jpeg", device="jpeg", plot=K_M_line, path="G:/POP_STRUCTURE_PUB/Figures", dpi="retina", width = 6.5, height = 5)
```


```{r, echo=FALSE}

(M_K<- ggplot(RG.M, aes(x=Month, y=K*100000, fill=Site)) + geom_boxplot(position=position_dodge(1)) + labs(y=NULL, x =NULL, title = "Male") + ylim(0, 2.5) + scale_fill_manual(values=c( "black", "grey")) + labs(fill = "Site") + theme(legend.position = c(0.8, 0.8), axis.text.x = element_text( size = 12), axis.text.y = element_text( size = 12), axis.title.x = element_text( size = 15), axis.title.y = element_text( size = 13), legend.text = element_text(size=13), legend.title = element_text(size=15)) + theme_classic()+ ggtitle("a)"))

```


## Female K


#### a. Normality tests (Shapiro Wilk) & Kruskal Wallis and dunn test post hoc multiple comparison

```{r, echo=FALSE}
shapiro.test(RG.F$K)
kruskal.test(K ~ Month, data = RG.F)
(K_F_Comp<-dunn_test(K ~ Month, data = RG.F, p.adjust.method = "holm"))

```


#### b. Summary statistics and monthly plot creation

```{r, echo=FALSE}
(F_K_Summary<-(summarySE(RG.F, measurevar="K", groupvars=c("Site", "Month"), na.rm = TRUE)))

(K_F_line<- ggplot(F_K_Summary, aes(y=K*100000, x=Month, colour=Site)) + ylim(1, 2) + geom_errorbar(aes(ymin=(K-se)*100000, ymax=(K+se)*100000), width=.1) + geom_point() + labs(title = "Female") + geom_line(aes(group=Site))+ theme_classic () + scale_colour_manual(values=c("grey", "black")) + theme(plot.margin = margin(t=0.5, l=0.5, r=0.5, b=0.5, unit = "cm"), text = element_text(size = 15), legend.position = "bottom") + labs(y=bquote("Fulton's Condition \n (Total Weight / Total" *Length^3~")*" * 10^5~")")))

ggsave("K_Lineplot_female.jpeg", device="jpeg", plot=K_F_line, path="G:/POP_STRUCTURE_PUB/Figures", dpi="retina", width = 6.5, height = 5)

```

```{r, echo=FALSE}

(F_K<- ggplot(RG.F, aes(x=Month, y=K*100000, fill=Site)) + geom_boxplot(position=position_dodge(1)) + labs( y=NULL, x = NULL, title = "Female")  + ylim(0, 2.5) + scale_fill_manual(values=c( "black", "grey ")) + labs(fill = "Site") + theme(legend.position = c(0.8, 0.8), axis.text.x = element_text( size = 12), axis.text.y = element_text( size = 12), axis.title.x = element_text( size = 15), axis.title.y = element_text( size = 13), legend.text = element_text(size=13), legend.title = element_text(size=15))+ theme_classic())
```


## M-F plot bind

```{r, echo=FALSE}
K_LINE<-ggarrange(K_M_line + rremove ("ylab"), K_F_line +  rremove ("ylab"), ncol = 1, nrow = 2, common.legend = TRUE, legend = "bottom" , hjust = c(-1, -0.7))

(K_LINE<- annotate_figure(K_LINE,left=text_grob("Fulton's Condition (K) \n (Total Weight/Total Length^3 * 100,000)", rot = 90, size = 15)))

ggsave("K_LINE.jpeg", device="jpeg", plot=K_LINE, path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures", dpi="retina", width = 7.48031, height = 9)
```





