---
title: "Back Calculated Length at Age"
author: "Anna Conklyn"
date: "3/8/2023"
output: html_document
---

# **Structure and Condition of an Established Round Goby Population in the Thousand Islands Region, St. Lawrence River**
## *Back Calculated Length at Age*

### Contents of Rmarkdown:
  > 1. Data setup, convert from one-fish-per-line to one-measurement-per-line
  > 2. Linear regression fitting of otolith radius vs. total length to aquire coefficients for back calculation
  > 3. Back-calculate total lengths at age
  > 4. Create summary tables and plots for predicted length-at-age
  
```{r setup, include=FALSE}
##Install packages

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
library(FSA)
library(nlstools)
library(magrittr)
library(dplyr)
library(plotrix)
library(nnet)
library(nls2)
library(ggplot2)
library(tidyr)
library(broom)
library(stargazer)
library(plyr)
library(gt)
library(gtsummary)
library(stringr)
library(lattice)
library(car)
library(multcomp)
library(readxl)
```


```{r, include=FALSE}
##Load data


getwd()
setwd("G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE")
RG <- read_excel("Pop_Strucutre_Dissected.xlsx")

RG<- filter(RG, Age != "NA")
```



#### 1. Data setup, convert from one-fish-per-line to one-measurement-per-line

```{r, include=FALSE}
RG$yearclass<-RG$Year-RG$Age

Male<- filter(RG, Sex== "M")
Female<- filter(RG, Sex=="F")
```


```{r, include=FALSE}
RG_long_M<- Male %>% pivot_longer(cols = c( 'RadAnnu0', 'RadAnnu1', 'RadAnnu2', 'RadAnnu3', 'RadAnnu4', 'RadAnnu5'), names_to = "AgeRad", values_to = "Rad")
RG_long_M$AgeRad<- str_remove_all(RG_long_M$AgeRad, "[RadAnnu]")
RG_long_M$AgeRad<- as.numeric(RG_long_M$AgeRad)

RG_long_F<- Female %>% pivot_longer(cols = c( 'RadAnnu0', 'RadAnnu1', 'RadAnnu2', 'RadAnnu3', 'RadAnnu4', 'RadAnnu5'), names_to = "AgeRad", values_to = "Rad")
RG_long_F$AgeRad<-str_remove_all(RG_long_F$AgeRad, "[RadAnnu]")
RG_long_F$AgeRad<- as.numeric(RG_long_F$AgeRad)
```


### 2. Linear regression fitting of otolith radius vs. total length to aquire coefficients for back calculation

#### Male

```{r, echo=FALSE}
BodyOR_M<- lm(log(Total.Length) ~ log(OR), data=Male) 
(BodyOR_M_Summary<- coef(BodyOR_M))
a_M<- coef(BodyOR_M)[2]
```

#### Female
```{r, echo=FALSE}
BodyOR_F<- lm(log(Total.Length) ~ log(OR), data=Female) 
(BodyOR_F_Summary<- coef(BodyOR_F))
a_F<- coef(BodyOR_M)[2]
```

### 3. Back-calculate total lengths at age

$$ TL_{n} = TL * (OR_{n}/OR)^a $$
TL_{n} is the total length at annulus n; TL, observed total length; OR_{n} otolith radius at annulus n; OR, otolith radius at capture; a, the slope of linear function between logarithm of total length and logarithm of total radius 

#### Male
```{r, echo=FALSE}
RG_long_M$Pred_len<- with(RG_long_M, (Total.Length)*(Rad/OR)^a_M)

ggplot(RG_long_M,aes(Pred_len)) + geom_histogram(aes(fill=Site)) + theme_classic() + scale_fill_manual(values = c("Rose" = "black", "Cobb" = "grey")) +  theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 15), legend.position = "bottom")
```

#### Female
```{r, echo=FALSE}
RG_long_F$Pred_len<- with(RG_long_F, (Total.Length)*(Rad/OR)^a_F)

ggplot(RG_long_F,aes(Pred_len)) + geom_histogram(aes(fill=Site)) + theme_classic() + scale_fill_manual(values = c("Rose" = "black", "Cobb" = "grey")) +  theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 15), legend.position = "bottom")
```











# Statistical differences 


```{r, echo=FALSE}

MaleBC_Site<- lm(AgeRad~ Pred_len * Site, data = RG_long_M)
(MaleBC_Site_Sum<- summary(MaleBC_Site))
residualPlots(MaleBC_Site)

(MaleBC_Site_regression<- tbl_regression(MaleBC_Site) %>%  
  add_significance_stars(hide_p = FALSE,
                           pattern = "{p.value}{stars}") %>%
  add_glance_table(
    include = c(nobs, r.squared)))


(MaleBC_table<-Anova(MaleBC_Site))
rownames(MaleBC_table)[rownames(MaleBC_table) == "AgeRad"] <- "Age"
rownames(MaleBC_table)[rownames(MaleBC_table) == "AgeRad:Site"] <- "Age:Site"
MaleBC_table$rownames<- rownames(MaleBC_table)

(MaleBC_table_print <- MaleBC_table %>% gt(rowname_col = "rownames") %>% fmt_number(columns = c("Sum Sq" , "F value"),  decimals = 2)  %>% tab_header( title = md("**Male**")))


```





```{r}
FemaleBC_Site<- lm(AgeRad ~ Pred_len*Site, data = RG_long_F)
(FemaleBC_Site_Sum<- summary(FemaleBC_Site))
residualPlots(FemaleBC_Site)
(FemaleBC_table<-Anova(FemaleBC_Site, type = 3))
rownames(FemaleBC_table)[rownames(FemaleBC_table) == "AgeRad"] <- "Age"
rownames(FemaleBC_table)[rownames(FemaleBC_table) == "AgeRad:Site"] <- "Age:Site"
FemaleBC_table$rownames<- rownames(FemaleBC_table)

(FemaleBC_Site_regression<- tbl_regression(FemaleBC_Site) %>%  
  add_significance_stars(hide_p = FALSE,
                           pattern = "{p.value}{stars}") %>%
  add_glance_table(
    include = c(nobs, r.squared)))



(FemaleBC_table_print <- FemaleBC_table %>% gt(rowname_col = "rownames") %>% fmt_number(columns = c("Sum Sq" , "F value"),  decimals = 2)  %>% tab_header( title = md("**Male**")))
```
```{r}
(BC_Regression_Table<- tbl_merge(
    tbls = list(MaleBC_Site_regression, FemaleBC_Site_regression),
    tab_spanner = c("**Male**", "**Female**")
  ) %>% as_gt())
 
BC_Regression_Table %>% gtsave(filename="BC_Test_Table.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures")
```




### Create summary tables and plots for predicted length-at-age

### F-test for equality of slopes for back-calculated length at age regression


```{r, echo=FALSE}

MaleBC_table$Sex<-"Male"
FemaleBC_table$Sex<-"Female"

FullBC_ANCOVA_table<- rbind(MaleBC_table, FemaleBC_table)

(FullBC_ANCOVA<- gt(FullBC_ANCOVA_table, rowname_col = "rownames", groupname_col = "Sex") %>% fmt_number(columns = c("Sum Sq" , "F value", "Pr(>F)"),  decimals = 2) %>% 
  tab_style(
    style = cell_text( weight = "bold", align = "left"),
    locations = cells_group()) %>% 
  tab_style(
    style = cell_text( weight = "bold", align = "left"),
    locations = cells_column_labels()
  ))
  
FullBC_ANCOVA %>% gtsave(filename = "BCANCOVA table.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures")

```



#### Rose Male Predictions

```{r, echo=FALSE}
RG_long_M_Rose<- filter(RG_long_M, Site=="Rose")
M_R_BC_Sum<- Summarize(Pred_len~AgeRad,data=RG_long_M_Rose,digits=2)
M_R_BC_Sum$Sex<- "Rose Male"

(Male_Rose_BC_table =  gt(M_R_BC_Sum[1:4], rowname_col = "AgeRad") %>% cols_hide(columns="n") %>% tab_header(title="Rose Male") %>% cols_label(
    nvalid = "n",
    mean = "mean") %>% cols_align(align = "left") %>%
  tab_stubhead(label = "Age") %>% tab_options(table.width = pct(50))%>% cols_width(everything() ~ px(100)))
```

#### Cobb Male Predictions

```{r, echo=FALSE}
RG_long_M_Cobb<- filter(RG_long_M, Site=="Cobb")
M_C_BC_Sum<- Summarize(Pred_len~AgeRad,data=RG_long_M_Cobb,digits=2)
M_C_BC_Sum$Sex<- "Cobb Male"


(Male_Cobb_BC_table =  gt(M_C_BC_Sum[1:4], rowname_col = "AgeRad") %>% cols_hide(columns="n") %>% tab_header(title="Cobb Male") %>% cols_label(
    nvalid = "n",
    mean = "mean") %>% cols_align(align = "left") %>%
  tab_stubhead(label = "Age") %>% tab_options(table.width = pct(50))%>% cols_width(everything() ~ px(100)))
```


## Cobb Females
```{r, echo=FALSE}
RG_long_F_Cobb<- filter(RG_long_F, Site=="Cobb")
F_C_BC_Sum<- Summarize(Pred_len~AgeRad,data=RG_long_F_Cobb,digits=2)
F_C_BC_Sum$Sex<- "Cobb Female"

(Female_Cobb_BC_table = gt(F_C_BC_Sum[1:5], rowname_col = "AgeRad") %>% cols_hide(columns="n") %>% tab_header(title="Cobb Female") %>% cols_label(
    nvalid = "n",
    mean = "mean") %>% cols_align(align = "left") %>%
  tab_stubhead(label = "Age") %>% tab_options(table.width = pct(50))%>% cols_width(everything() ~ px(100)))
```

## Rose Females
```{r, echo=FALSE}
RG_long_F_Rose<- filter(RG_long_F, Site=="Rose")
F_R_BC_Sum<- Summarize(Pred_len~AgeRad,data=RG_long_F_Rose,digits=2)
F_R_BC_Sum$Sex<- "Rose Female"

(Female_Rose_BC_table = gt(F_R_BC_Sum[1:5], rowname_col = "AgeRad") %>% cols_hide(columns="n") %>% tab_header(title="Cobb Female") %>% cols_label(
    nvalid = "n",
    mean = "mean") %>% cols_align(align = "left") %>%
  tab_stubhead(label = "Age") %>% tab_options(table.width = pct(50))%>% cols_width(everything() ~ px(100)))
```

```{r}

FullBC<- rbind(M_R_BC_Sum, M_C_BC_Sum, F_R_BC_Sum, F_C_BC_Sum)
FullBC <- FullBC 

(FullBC_Table= gt(FullBC, rowname_col = "AgeRad", groupname_col = "Sex") %>% cols_hide(columns=c("n", "min", "Q1", "median", "Q3", "max")) %>% tab_header(title= md("**Mean back-calculated length-at-age**")) %>% cols_label(
    nvalid = md("**n**"),
    mean = md("**Mean**"), sd = md("**SD**")) %>% cols_align(align = "left") %>%
  tab_stubhead(label = md("**Age**")) %>% tab_options(table.width = pct(50)) %>% cols_width(everything() ~ px(100)) )

 
FullBC_Table %>% gtsave(filename="BC_Table.png", path="G:/POP_STRUCTURE_PUB/Figures")
```



## Jitter plot for back calculated length-at-age

``` {r}
(M_BC_jitter<- ggplot(RG_long_M, aes(AgeRad, Pred_len)) + geom_jitter(aes(colour=Site)) +  facet_wrap(vars(Age)) + ylim (0,300) + xlim(0,5) + ylab ("Back-Calculated Length (mm)") + xlab ("Age (Jittered)") + labs(title = "Male") + theme_classic () + scale_color_manual(values = c("Rose" = "black", "Cobb" = "grey")) +  theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 15), legend.position = "bottom"))

M_BC_jitter %>% ggsave(filename="Male BC_jitter.png", path="G:/POP_STRUCTURE_PUB/Figures", dpi="retina", width = 7.48031, height = 5)

```

```{r}
(F_BC_jitter<- ggplot(RG_long_F, aes(AgeRad, Pred_len)) + geom_jitter(aes(colour=Site)) + facet_wrap(vars(Age)) + ylim (0,300) + xlim(0,5) + ylab ("Back-Calculated Length (mm)") + xlab ("Age (Jittered)") + labs(title = "Female") + theme_classic () + scale_color_manual(values = c("Rose" = "black", "Cobb" = "grey")) +  theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 15), legend.position = "bottom"))

 F_BC_jitter %>% ggsave(filename="Female BC_jitter.png", path="G:/POP_STRUCTURE_PUB/Figures")

```

```{r}
(BCplot <- ggarrange(M_BC_jitter+ rremove ("ylab"), F_BC_jitter+ rremove ("ylab"),
                    nrow = 2, common.legend = TRUE, legend="bottom"))

(BCplot<- annotate_figure(BCplot,left=text_grob("Back-Calculated Total Length (mm)", rot = 90, size = 15)))

ggsave("BCplot.jpeg", device="jpeg", plot=BCplot, path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures", dpi="retina", width = 7.48031, height = 9)
```


