---
title: "Maturity"
author: "Anna Conklyn"
date: "2023-11-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(gtExtras)
library(FSA)
library(car)
library(plyr)
library(tidyr)
library(dplyr)
library(magrittr)
library(ggplot2)
library(ggpubr)
library(sjPlot)
library(gt)
library(cowplot)
```


# **Structure and Condition of an Established Round Goby Population in the Thousand Islands Region, St. Lawrence River**
## *Maturity analysis*

### Contents of Rmarkdown:
  > 1.   Maturity at length
      > a. Logistic regression model fitting
      > b. Compute age-at-50% and 90% maturity
      > c. Statistical comparisons
  
  > 2. Maturity at age
      > a. Logistic regression model fitting
      > b. Compute age-at-50% and 90% maturity
      > c. Statistical comparisons
      
  > 3. Summary figures
  
```{r, include=FALSE, data setup}
setwd("G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE")
AllRG <- read_excel("Pop_Strucutre_Dissected.xlsx") 

AllRG %<>% mutate(LCat=lencat(Total.Length, w=10)) %<>% filter(!is.na(AllRG$Age)) 

as.factor(AllRG$Month)
table(AllRG$Month)

spawning<-(c("July", "June", "May"))

AllRG <- AllRG %>% filter(Month %in% spawning)

AllRG %>% group_by(Site)
AllRG$Maturity<- as.factor(AllRG$Maturity)
AllRG$Maturity1<-ifelse(AllRG$Maturity=="Mature",1,0)

RG.C<- AllRG %>% filter(Site == "Cobb")
RG.R<- AllRG %>% filter(Site == "Rose")

```


# Maturity at length

### Logistic model fitting
```{r, maturity at length}
glm_sites_length<- glm(Maturity~Total.Length*Site, data = AllRG, family=binomial)
Length_Chi<- drop1(glm_sites_length,~.,test="Chisq")
print(Length_Chi)

Site_Length_Maturity_table<- tbl_regression(glm_sites_length) %>%
  add_glance_table(
    include = c(nobs))  %>%  
  add_significance_stars(hide_p = FALSE,
                           pattern = "{p.value}{stars}")
```

### Compute length-at- 50% and 90% maturity for each site

$$ x = \frac {log (\frac{p}{1-p}) - \alpha} {\beta_1} $$

```{r, echo=FALSE}

lrPerc <- function(cf,p) (log(p/(1-p))-cf[1])/cf[2]

``` 

#### Cobb
```{r, echo=FALSE}
glm_C <- glm(Maturity1~Total.Length,data=RG.C,family=binomial)
coef(glm_C)

(C_L50 <- lrPerc(coef(glm_C),0.5))
C_L50$Site<-"Cobb"
C_L50$Percentage<-"50%"

(C_L90 <- lrPerc(coef(glm_C),0.9))
C_L90$Site<-"Cobb"
C_L90$Percentage<-"90%"


```

### Rose
```{r, echo=FALSE}

glm_R <- glm(Maturity1~Total.Length,data=RG.R,family=binomial)

(R_L50 <- lrPerc(coef(glm_R),0.5))
R_L50$Site<-"Rose"
R_L50$Percentage<-"50%"

(R_L90 <- lrPerc(coef(glm_R),0.9))
R_L90$Site<-"Rose"
R_L90$Percentage<-"90%"

```

```{r}
(Maturity_LPred<- rbind(C_L50, C_L90, R_L50, R_L90))
Maturity_LPred<- gtsummary::as_tibble(Maturity_LPred)
Maturity_LPred$`(Intercept)` %<>% as.numeric(Maturity_LPred$`(Intercept)`) %<>% round(., digits = 2)

```








# Maturity at age

### Logistic model fitting
```{r, maturity at age}

AllRG %<>% filter(!is.na(AllRG$Age))
glm_sites_age<- glm(Maturity~Age*Site, data = AllRG, family=binomial)
Age_Chi<- drop1(glm_sites_age,~.,test="Chisq")
print(Age_Chi)

(Site_Age_Maturity_table<- tbl_regression(glm_sites_age) %>% add_glance_table(
    include = c(nobs)) %>%  
  add_significance_stars(hide_p = FALSE,
                           pattern = "{p.value}{stars}"))
```

#### Cobb
```{r, echo=FALSE}

glm_C <- glm(Maturity1~Age,data=RG.C,family=binomial)
coef(glm_C)

(C_L50 <- lrPerc(coef(glm_C),0.5))
C_L50$Site<-"Cobb"
C_L50$Percentage<-"50%"
(C_L90 <- lrPerc(coef(glm_C),0.9))
C_L90$Site<-"Cobb"
C_L90$Percentage<-"90%"

```

### Rose
```{r, echo=FALSE}

glm_R <- glm(Maturity1~Age,data=RG.R,family=binomial)
coef(glm_C)

(R_L50 <- lrPerc(coef(glm_R),0.5))
R_L50$Site<-"Rose"
R_L50$Percentage<-"50%"
(R_L90 <- lrPerc(coef(glm_R),0.9))
R_L90$Site<-"Rose"
R_L90$Percentage<-"90%"

```

```{r}
(Maturity_APred<- rbind(C_L50, C_L90, R_L50, R_L90))
Maturity_APred<- gtsummary::as_tibble(Maturity_APred)
Maturity_APred$`(Intercept)` %<>% as.numeric(Maturity_APred$`(Intercept)`) %<>% round(., digits = 2)

Age_Maturity_Perc<- gt(Maturity_APred, groupname_col = "Site")
```


# Summary figures

```{r, summary table}

(Maturity_Table<- tbl_merge(
    tbls = list(Site_Length_Maturity_table, Site_Age_Maturity_table),
    tab_spanner = c("**Maturity-at-Length**", "**Maturity-at-Age**")
  ) %>% as_gt())
 
 Maturity_Table %>% gtsave(filename="Maturity_Table.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures")
 
```

```{r}
(Maturity_length_plot<- ggplot(AllRG, aes(x=Total.Length, y=Maturity1, color=Site)) + geom_point(aes(color=Site)) + labs(title = "A") +
  stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + theme_classic () + scale_colour_manual(values=c("grey", "black")) + labs(x= "Total length (mm)", y="Probability of maturity")+ theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 15), legend.position = "bottom"))


(Maturity_age_plot<- ggplot(AllRG, aes(x=Age, y=Maturity1, color=Site)) + geom_point(aes(color=Site)) + labs(title = "B") + 
  stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + theme_classic () + scale_colour_manual(values=c("grey", "black")) + labs(x= "Age", y="Probability of maturity")+ theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 15), legend.position = "bottom"))


(Maturity<- ggarrange(Maturity_length_plot + rremove("ylab"), Maturity_age_plot + rremove("ylab"), ncol = 1, nrow = 2, common.legend = TRUE,  hjust = c(-2,-.5), vjust=c(1.2,1.2), legend = "bottom"))
(Maturity<- annotate_figure(Maturity,left=text_grob("Probability of Maturity", rot = 90, size = 15)))
ggsave("Maturity_plot.jpeg", device="jpeg", plot=Maturity, path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures", dpi="retina", width = 7.48031, height = 9)

```

```{r}
library(gtExtras) 

(Maturity_LPerc<-gt(Maturity_LPred, groupname_col = "Site", rowname_col = "Percentage") %>% tab_header(title = "Length at maturity")  %>%
  cols_label(  "(Intercept)" = md("**Estimate**"))%>% 
  tab_style(
    style = cell_text( weight = "bold", align = "left"),
    locations = cells_group()) %>% 
  tab_style(
    style = cell_text( weight = "bold", align = "left"),
    locations = cells_title() 
    ))
Maturity_LPerc %>% gtsave(filename="Maturity_Length_Table.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE")


(Maturity_APerc<-gt(Maturity_APred, groupname_col = "Site", rowname_col = "Percentage") %>% tab_header(title = "Age at maturity") %>%
  cols_label(  "(Intercept)" = md("**Estimate**")) %>% 
  tab_style(
    style = cell_text( weight = "bold", align = "left"),
    locations = cells_group()) %>% 
  tab_style(
    style = cell_text( weight = "bold", align = "left"),
    locations = cells_title()
  ))
Maturity_APerc %>% gtsave(filename="Maturity_Age_Table.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE")

(Maturity_LPerc_image <- ggdraw() + draw_image("Maturity_Length_Table.png"))
(Maturity_APerc_image <- ggdraw() + draw_image("Maturity_Age_Table.png"))

(Maturity_predictions<- ggarrange(Maturity_LPerc_image, Maturity_APerc_image, ncol = 2))

ggsave("Maturity_predictions.jpeg", device="jpeg", plot=Maturity_predictions, path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures", dpi="retina", width = 7.48031, height = 9)
```

