---
title: "Mortality"
author: "Anna Conklyn"
date: "11/17/2022"
output:
  html_document: default
  word_document: default
---

# **Structure and Condition of an Established Round Goby Population in the Thousand Islands Region, St. Lawrence River**

## *Mortality*

### Contents of Rmarkdown:

    > 1. Catch-curve and Chapman-Robson plot creation
    > 2. Linear regression fitting and survival estimate
    > 3. Statistical comparisons
    > 4. Publishable plot creation

```{r setup, include=FALSE}
library(FSA)
library(nlstools)
library(ggplot2)
library(dplyr)
library(magrittr)
library(plotrix)
library(nnet)
library(nls2)
library(ggpubr)
library(rlang)
library(readxl)
library(ggpubr)
library(sjPlot)
library(gt)
library(grid)
library(patchwork)
library(kableExtra)
library(gtsummary)
library(webshot)
library(gtExtras)
```

```{r, include=FALSE}
getwd()
setwd("G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE")
RG <- read_excel("Pop_Strucutre_Dissected.xlsx")
options(digits = 3)
```

```{r, include=FALSE}
RG %<>% filter(!is.na(RG$Age))
RG %<>% filter(Sex != "I")

RGCounts_site<- RG %>% count(Age, Site, sort = TRUE)
RGCounts_sex<- RG %>% count(Age, Sex, sort = TRUE)

RGCounts<- RG %>% count(Age, sort = TRUE)


Cobb<- filter(RG, Site=="Cobb")
CCounts<- Cobb %>% count(Age)
Rose<- filter(RG, Site== "Rose")
RCounts<- Rose %>% count(Age)

Male<- filter(RG, Sex== "M")
Male %>% group_by(Site)
MCounts<- Male %>% count(Age)
CMale<- filter(Male, Site== "Cobb")
RMale<- filter(Male, Site== "Rose")

Female<- filter(RG, Sex== "F")
Female %>% group_by(Site)
(FCounts<- Female %>% count(Age))
CFemale<- filter(Female, Site== "Cobb")
RFemale<- filter(Female, Site== "Rose")


(CMCounts<- CMale %>% count(Age))
(RMCounts<- RMale %>% count(Age))
(CFCounts<- CFemale %>% count(Age))
(RFCounts<- RFemale %>% count(Age))
```

#### A=Annual mortality

#### S= Annual Survival

#### Z=instantaneous mortality rate

$$\ A+S=1 $$ $$\ A=\frac{C_{t}-C_{t+1}}{C_{t}}=1-\frac{C_{t+1}}{C_{t}}$$ $$\ Z=log(C_{t})-log(C_{t+1})$$ $$\ A=1-e^{-Z} $$

## 1. Chapman-Robson Metric estimation and catch curve

Estimates Annual Survival Rate using:

$$\widehat{S}= \frac{T}{n+T-1}=\frac{\overline{T}}{1+\overline{T}-\frac{1}{n}}$$

### All gobies combined
```{r, echo=FALSE}
RG_cr <- chapmanRobson(n~Age,data=RGCounts,ages2use=3:5)
Measure<- c("S", "Z")
AllRG_CR<-cbind(summary(RG_cr, digits=2), confint(RG_cr, digits=2))
(AllRG_CR<- data.frame(Measure, AllRG_CR))


RG_CC<- catchCurve(n~Age, data=RGCounts, ages2use =3:5, weighted = TRUE)
plot(RG_CC, pos.est = "bottomleft")
Measure<- c("Z", "A")
(RG_CC<- cbind(summary(RG_CC, digits=2), confint(RG_CC, digits=2)))
(AllRG_CC<- data.frame(Measure, RG_CC))
AllRG_CC<- AllRG_CC[2,]
AllRG_CC<- AllRG_CC[-c(4,5)]

(AllRG_Mortality_table<- rbind(AllRG_CR, AllRG_CC))

```

### Sexes combined

#### Cobb

```{r, echo=FALSE}
Ccr <- chapmanRobson(n~Age,data=CCounts,ages2use=2:4)
Group<- "Cobb"
Measure<- c("S", "Z")
Cobb_CR<-cbind(summary(Ccr, digits=2), confint(Ccr, digits=2))
(Cobb_CR<- data.frame(Measure, Cobb_CR, Group))


CCC<- catchCurve(n~Age, data=CCounts, ages2use = 2:4, weighted = TRUE)
(Cobb_CC<- cbind(summary(CCC, digits=2), confint(CCC, digits=2)))
plot(CCC, pos.est = "bottomleft")
```

#### Rose

```{r, echo=FALSE}
Rcr <- chapmanRobson(n~Age,data=RCounts,ages2use=1:5)
Group<- "Rose"
Rose_CR<-cbind(summary(Rcr, digits=2), confint(Rcr, digits=2))
(Rose_CR<- data.frame(Measure, Rose_CR, Group))

RCC<- catchCurve(n~Age, data=RCounts, ages2use = 3:5, weighted = TRUE)
(Rose_CC<- cbind(summary(RCC, digits=2), confint(RCC, digits=2)))
plot(RCC, pos.est = "bottomleft")

```

### Sites Combined

#### Male

```{r, echo=FALSE}
Mcr <- chapmanRobson(n~Age,data=MCounts,ages2use=3:5)
Group<- "Male"
Male_CR<-as_tibble(cbind(summary(Mcr, digits=2), confint(Mcr, digits=2)))
(Male_CR<- data.frame(Measure, Male_CR, Group))

MCC<- catchCurve(n~Age, data=MCounts, ages2use = 3:5, weighted = TRUE)
(Male_CC<- cbind(summary(MCC), confint(MCC)))
plot(MCC, pos.est = "bottomleft")

```

#### Female

```{r, echo=FALSE}
Fcr <- chapmanRobson(n~Age,data=FCounts,ages2use=2:5)
Group<- "Female"
Female_CR<-as_tibble(cbind(summary(Fcr,digits=3),confint(Fcr,digits=3)))
(Female_CR<- data.frame(Measure, Female_CR, Group))

FCC<- catchCurve(n~Age, data=FCounts, ages2use = 2:5, weighted = TRUE)
(Female_CC<- cbind(summary(FCC), confint(FCC)))
plot(FCC, pos.est = "bottomleft")

```


## 2. Linear regression fitting, survival estimate, and statistical comparisons 

### Site

```{r, echo=FALSE}
RGCounts_site_2<- filter(RGCounts_site, Age>=3)
RGCounts_site_2$lnct<- log(RGCounts_site_2$n)

lmCC_site<- lm(Age~lnct*Site, data=RGCounts_site_2)
summary(lmCC_site)
anova(lmCC_site)

coef<- coef(lmCCs_site)
(zhat_Cobb <- (-1*coef[["Age"]]))
(zhat_Rose <- (-1*coef[["Age"]]+coef[["Age:SiteRose"]]))


(Site_Mortality_Table<- tbl_regression(lmCC_site) %>%  
  add_significance_stars(hide_p = FALSE,
                           pattern = "{p.value}{stars}") %>%
  add_glance_table(
    include = c(nobs, r.squared)))

```

### Sex

```{r, echo=FALSE}
RGCounts_sex_2<- filter(RGCounts_sex, Age>=3)
RGCounts_sex_2$lnct<- log(RGCounts_sex_2$n)

lmCC_sex<- lm(Age~lnct*Sex, data=RGCounts_sex_2)
summary(lmCC_sex)
anova(lmCC_sex)

coef<- coef(lmCC_sex)
(zhat_Female <- -1*coef[["Age"]])
(zhat_Male <- -1*coef[["Age"]]+coef[["Age:SexM"]])

(Sex_Mortality_Table<- tbl_regression(lmCC_sex) %>%  
  add_significance_stars(hide_p = FALSE,
                           pattern = "{p.value}{stars}") %>%
  add_glance_table(
    include = c(nobs, r.squared)))

```


## 4. Plot and table creation

### Plots
#### Sites

```{r, echo=FALSE}

(Site_plot<-ggplot(data=RGCounts_site, aes(x=Age, y=log(n))) + geom_point(aes(colour=Site)) + ylim(0,5) + stat_smooth( method = lm, se=FALSE, aes(colour=Site), data=RGCounts_site_2) + labs(title="A") + scale_colour_manual(values=c("black", "grey")) + labs(x= "Age", y="log(Catch [n])")+ theme_classic() + theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 15), legend.position = "bottom"))

ggsave("Mortality_site.jpeg", device="jpeg", plot=Site_plot, path="G:/POP_STRUCTURE_PUB/Figures", dpi="retina", width = 7.48031, height = 5)
```

#### Sexes


```{r, echo=FALSE}

(Sex_plot<-ggplot(data=RGCounts_sex, aes(x=Age, y=log(n))) + geom_point(aes(colour=Sex)) + ylim(0,5) + stat_smooth( method = lm, se=FALSE, aes(colour=Sex), data = RGCounts_sex_2) + labs(title="B") + scale_colour_manual(values=c("black", "grey")) + labs(x= "Age", y="log(Catch [n])")+ theme_classic() + theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 15), legend.position = "bottom"))

ggsave("Mortality_sex.jpeg", device="jpeg", plot=Sex_plot, path="G:/POP_STRUCTURE_PUB/Figures", dpi="retina", width = 7.48031, height = 5)
```

#All gobies combined

```{r, echo=FALSE}

RGCounts$lnct<- log(RGCounts$n)
RGCounts_3<- filter(RGCounts, Age>=3)


(ALL_Mortality_plot<-ggplot(data=RGCounts, aes(x=Age, y=log(n))) + geom_point() + ylim(0,7) + labs(x= "Age", y="log(Catch [n])") + stat_smooth(method = lm, se=FALSE, data = RGCounts_3, colour="black") + theme_classic() + theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 15)))

ggsave("Mortality_all.jpeg", device="jpeg", plot=ALL_Mortality_plot, path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures", dpi="retina", width = 7.48031, height = 5)
```


```{r, include=FALSE}
(Mortality_plot<- ggarrange(Site_plot+ rremove ("ylab"), Sex_plot + rremove ("ylab"), ncol = 1, nrow = 2, common.legend = FALSE, align = "h", hjust = c(0, 1.5), vjust=c(0.5, 0.5), legend = "bottom"))
(Mortality_plot<- annotate_figure(Mortality_plot,left=text_grob("log(Catch [n])", rot = 90, size = 15)))

ggsave("Mortality_Plot.jpeg", device="jpeg", plot=Mortality_plot, path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures", dpi="retina", width = 7.48031, height = 9)
```


## Linear model F-tests

```{r}

(Mortality_Table<- tbl_merge(
    tbls = list(Site_Mortality_Table, Sex_Mortality_Table),
    tab_spanner = c("**Mortality by Site**", "**Mortality by Sex**")
  ) %>% as_gt())
 
 Mortality_Table %>% gtsave(filename="Mortality_Table.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures")
```


### Chapman Robson metrics tables

## All RG table
```{r,echo=FALSE}


(Mortality_table_all = gt(AllRG_Mortality_table, rowname_col = "Measure") %>% tab_stubhead(label = md("**Metric**")) %>%
  cols_merge(columns = c(X95..LCI, X95..UCI), pattern = "({1},{2})" ) %>% tab_options(table.width = pct(60)) %>%
    cols_label(Estimate=md("**Estimate**"), Std..Error=md("**Standard Error**"), X95..LCI= md("**95% CI**")) %>%  cols_width(everything() ~ px(100)) %>%
 cols_align(align="left", columns = everything()) )
  


Mortality_table_all %>% gtsave(filename="Mortality_table_all.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures")

```

#Sites separated table

```{r,echo=FALSE}
Site_table<- rbind(Cobb_CR, Rose_CR)
Site_table<-data.frame(lapply(Site_table, function(x) if(is.numeric(x)) round(x, 3) else x))



(Site_table = gt(Site_table, rowname_col = "Measure", groupname_col = "Group") %>% tab_header(title=md("**Mortality by Site**")) %>% tab_stubhead(label = md("**Metric**")) %>%
  cols_merge(columns = c(X95..LCI, X95..UCI), pattern = "({1},{2})" ) %>% tab_options(table.width = pct(60)) %>%
    cols_label(Estimate=md("**Estimate**"), Std..Error=md("**Standard Error**"), X95..LCI= md("**95% CI**")) %>%  cols_width(everything() ~ px(100)) %>%
  tab_row_group(
    group = md("**Cobb**"),
    rows = 1:2
  )  %>% tab_row_group(
    group = md("**Rose**"),
    rows = 3:4
  ) %>% cols_align(align="left", columns = everything()) )
  


Site_table %>% gtsave(filename="Site_Mortality_table.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE")

```


```{r, echo=FALSE}
Sex_table<- rbind(Female_CR, Male_CR)
Sex_table<-data.frame(lapply(Sex_table, function(x) if(is.numeric(x)) round(x, 3) else x))

(Sex_table= gt(Sex_table, rowname_col = "Measure", groupname_col = "Group") %>% tab_header(title=md("**Mortality by Sex**")) %>% tab_stubhead(label = md("**Metric**")) %>% cols_merge(columns = c(X95..LCI, X95..UCI), pattern = "({1},{2})" ) %>% tab_options(table.width = pct(60))  %>% cols_label(Estimate=md("**Estimate**"), Std..Error=md("**Standard Error**"), X95..LCI= md("**95% CI**")) %>%  cols_width(everything() ~ px(100)) %>%
  tab_row_group(
    group = md("**Female**"),
    rows = 1:2
  )  %>% tab_row_group(
    group = md("**Male**"),
    rows = 3:4
  ) %>% cols_align(align="left", columns = everything()) )

Sex_table %>% gtsave(filename="Sex_Mortality_table.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE")

```


```{r, include=FALSE}
library(cowplot)
library(patchwork)

(Sex_image <- ggdraw() + draw_image("Sex_Mortality_table.png"))
(Site_image <- ggdraw() + draw_image("Site_Mortality_table.png"))

(Chapman_Metrics<- ggarrange(Sex_image, Site_image, nrow = 2))

ggsave("Chapman_Metrics.jpeg", device="jpeg", plot=Chapman_Metrics, path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures", dpi="retina", width = 7.48031, height = 9)
```


