---
title: "Size structure and sex ratio"
author: "Anna Conklyn"
date: "2/5/2022"
output:
  pdf_document: default
  word_document: default
  html_document: default
---


# **Structure and Condition of an Established Round Goby Population in the Thousand Islands Region, St. Lawrence River**
## *Size structure and sex ratio*

### Contents of Rmarkdown:
  > 1. Summary tables of lengths by sex and site
  > 2. Length Frequency visualization and statistical analysis
  > 3. Sex ratio calculation, visualization and statistical analysis


```{r setup, include=FALSE}
library(FSA)
library(nlstools)
library(magrittr)
library(dplyr)
library(plotrix)
library(nnet)
library(nls2)
library(ggplot2)
library(readxl)
library(tables)
library(lubridate)
library(Matching)
library(ggpubr)
library(car)
library(tidyr)
library(gt)
library(gtExtras)
library(gtsummary)

knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

```



```{r, include=FALSE}
##Setup and data partitioning 

getwd()
setwd("G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE")
RG_Dissected <- read_excel("Pop_Strucutre_Dissected.xlsx")
RG_Dissected<- RG_Dissected %>% dplyr:: select("Date", "Total.Mass", "Total.Length", "Sex", "Site", "Month", "Year")


RG_NotDissected <- read_excel("Pop_Strucutre_NotDissected.xlsx")
RG_NotDissected<- RG_NotDissected %>% dplyr:: select("Date", "Total.Mass", "Total.Length", "Sex", "Site", "Month", "Year")


AllRG<- rbind(RG_Dissected, RG_NotDissected)

as.factor(AllRG$Month)
AllRG$Month<- factor(AllRG$Month, levels = c("May", "June", "July", "August", "September", "October"))

AllRG %<>% group_by(Month, Year, Site, Sex)
AllRG %<>% mutate(lcat5=lencat(Total.Length, w=5), lcat10=lencat(Total.Length, w=10), logLength=log(Total.Length))

Males<- AllRG %>% filter(Sex=="M")
Females<- AllRG %>% filter(Sex=="F")

Cobb<- AllRG %>% filter(Site=="Cobb")
Cobb_M<- Cobb %>% filter(Sex=="M")
Cobb_F<- Cobb %>% filter(Sex=="F")
Cobb_I<- Cobb %>% filter(Sex=="I")

Rose<- AllRG %>% filter(Site=="Rose")
Rose_M<- Rose %>% filter(Sex=="M")
Rose_F<- Rose %>% filter(Sex=="F")
Rose_I<- Rose %>% filter(Sex=="I")
```


## 1. Summary tables of lengths by sex, site, and month

```{r, echo = FALSE}

Cobb_M_Summary1<- Cobb_M %>% tbl_summary(by = Month, include = Total.Length, missing = "no", type = all_continuous() ~ "continuous2", statistic =  all_continuous() ~ c("{N_nonmiss}",  "{min}, {max}", "{mean}({sd})" )) %>% modify_header(all_stat_cols() ~ "**{level}**")

Cobb_F_Summary1<- Cobb_F %>% tbl_summary(by = Month, include = Total.Length, missing = "no", type = all_continuous() ~ "continuous2", statistic =  all_continuous() ~ c("{N_nonmiss}",  "{min}, {max}", "{mean}({sd})" )) %>% modify_header(all_stat_cols() ~ "**{level}**")

Cobb_I_Summary1<- Cobb_I %>% tbl_summary(by = Month, include = Total.Length, missing = "no", type = all_continuous() ~ "continuous2", statistic =  all_continuous() ~ c("{N_nonmiss}",  "{min}, {max}", "{mean}({sd})"))%>% modify_header(all_stat_cols() ~ "**{level}**")

(Cobb_Summary<- tbl_stack(tbls=list(Cobb_M_Summary1, Cobb_F_Summary1, Cobb_I_Summary1), group_header =c("Male", "Female", "Immature")) %>% as_gt() %>%
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups(groups = everything())
  ))

Cobb_Summary %>% gtsave(filename="Cobb_Length_Summary.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures")
                              
```



```{r, echo = FALSE,}

(Rose_M_Summary1<- Rose_M %>% tbl_summary(by = Month, include = Total.Length, missing = "no", type = all_continuous() ~ "continuous2", statistic =  all_continuous() ~ c("{N_nonmiss}", "{min}, {max}", "{mean}({sd})" )) %>% modify_header(all_stat_cols() ~ "**{level}**"))

(Rose_F_Summary1<- Rose_F %>% tbl_summary(by = Month, include = Total.Length, missing = "no", type = all_continuous() ~ "continuous2", statistic =  all_continuous() ~ c("{N_nonmiss}", "{min}, {max}", "{mean}({sd})" )) %>% modify_header(all_stat_cols() ~ "**{level}**"))

(Rose_I_Summary1<- Rose_I %>% tbl_summary(by = Month, include = Total.Length, missing = "no", type = all_continuous() ~ "continuous2", statistic =  all_continuous() ~ c("{N_nonmiss}", "{min}, {max}", "{mean}({sd})" )) %>% modify_header(all_stat_cols() ~ "**{level}**"))

(Rose_Summary<- tbl_stack(tbls=list(Rose_M_Summary1, Rose_F_Summary1, Rose_I_Summary1), group_header =c("Male", "Female", "Immature"))  %>% as_gt() %>%
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_row_groups(groups = everything())
  ) )

Rose_Summary %>% gtsave(filename="Rose_Length_Summary.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures")


```

## 2. Length Frequency visualization and statistical analysis

### Histogram of round goby length frequencies

```{r, echo=FALSE, message=FALSE}

library(ggplot2)
ggplot(AllRG, aes(x=Total.Length, fill=Sex)) + geom_histogram(binwidth = 10, color="black") + facet_grid(cols = vars(AllRG$Site)) + labs(title = "Histogram of round goby lengths and sexes", subtitle = "Thousand Islands, 2018-2019", x = "Total Length (mm)",  y = "Count") + scale_fill_grey() + theme_classic ()

(Cobb_length_plot<-ggplot(Cobb, aes(x=Total.Length, fill=Sex)) + geom_histogram(binwidth = 10, color="black") + labs(x = "Total Length (mm)",  y = "Count", title="Cobb") + scale_fill_grey() + theme_classic ()+ scale_fill_grey() + theme_classic () + theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 15), legend.position = "bottom") + ylim(0,250)+ xlim(0,250))
ggsave("Cobb_lengthCounts.jpeg", device="jpeg", plot=Cobb_length_plot, path="G:/POP_STRUCTURE_PUB/Figures", dpi="retina", width = 7.48031, height = 5)

(Rose_length_plot<-ggplot(Rose, aes(x=Total.Length, fill=Sex)) + geom_histogram(binwidth = 10, color="black") + labs(x = "Total Length (mm)",  y = "Count", title="Rose") + scale_fill_grey() + theme_classic ()+ scale_fill_grey() + theme_classic () + theme(plot.margin = margin(t=0.1, l=0.1, r=0.1, b=0.1, unit = "cm"), text = element_text(size = 15), legend.position = "bottom")+ ylim(0,250)+ xlim(0,250))
ggsave("Rose_lengthCounts.jpeg", device="jpeg", plot=Rose_length_plot, path="G:/POP_STRUCTURE_PUB/Figures", dpi="retina", width = 7.48031, height = 5)

Length_Histogram<- ggarrange(Cobb_length_plot + rremove ("ylab") , Rose_length_plot + rremove ("ylab"),
                    ncol = 1, nrow = 2, common.legend = TRUE, legend="bottom")

(Length_Histogram<-annotate_figure(Length_Histogram,left=text_grob("Count", rot = 90, size = 15)))

 
 ggsave("Length_Histogram.jpeg", device="jpeg", plot=Length_Histogram, path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures", dpi="retina", width = 7.48031, height = 9)
```

### Shapiro-Wilk normality test and qqplots

#### Males
##### Rose
```{r, echo= FALSE}
##Males
shapiro.test(Rose_M$Total.Length)
qqPlot(Rose_M$Total.Length)
```

##### Cobb
```{r, echo= FALSE}
shapiro.test(Cobb_M$Total.Length)
qqPlot(Cobb_M$Total.Length)
```


#### Females
##### Rose
```{r, echo= FALSE}
##Females
shapiro.test(Rose_F$Total.Length)
qqPlot(Rose_F$Total.Length)
```
##### Cobb
```{r, echo= FALSE}
shapiro.test(Cobb_F$Total.Length)
qqPlot(Cobb_F$Total.Length)
```



### ECDF plot, Kolmogrov-Smirnov test and bootstrapped K-S test 
  to determine whether the ECDF (empirical cumulative distribution function) are the same between two groups and can detect differences in the location (e.g., median), dispersion (e.g., variance), and shape of the distributions

#### Males
```{r, echo= FALSE}
##Male K-S
ks.test(Rose_M$Total.Length, Cobb_M$Total.Length)
ks.boot(Rose_M$Total.Length, Cobb_M$Total.Length)

clr <- c("black","gray50")
plot(ecdf(Rose_M$Total.Length),xlab="Total Length (mm)", ylab = "Fn(x)",
do.points=FALSE,verticals=TRUE,main="",col.01line=NULL)
plot(ecdf(Cobb_M$Total.Length),add=TRUE,do.points=FALSE,
verticals=TRUE,col=clr[2],col.01line=NULL)
legend("bottomright",c("Rose","Cobb"),col=clr,lty=1,
bty="n",cex=0.75)
title(main="Male RG ECDF")
```


#### Females
```{r, echo= FALSE}
##Female K-S
ks.test(Rose_F$Total.Length, Cobb_F$Total.Length)
ks.boot(Rose_F$Total.Length, Cobb_F$Total.Length)

clr <- c("black","gray50")
plot(ecdf(Rose_F$Total.Length),xlab="Total Length (mm)", ylab = "Fn(x)",
do.points=FALSE,verticals=TRUE,main="",col.01line=NULL)
plot(ecdf(Cobb_F$Total.Length),add=TRUE,do.points=FALSE,
verticals=TRUE,col=clr[2],col.01line=NULL)
legend("bottomright",c("Rose","Cobb"),col=clr,lty=1,
bty="n",cex=0.75)
title(main="Female RG ECDF")
```



### Chi Square test 
  to detect differences in length frequencies (5mm classes) between sites and months

##### Males
```{r, echo=FALSE}
RG_M_Site_xtab<- xtabs(~Site+lcat5, data=Males)
chisq.test(RG_M_Site_xtab)

RG_M_Month_xtab<- xtabs(~Month+lcat5, data=Males)
chisq.test(RG_M_Month_xtab)
```

##### Females
```{r, echo=FALSE}
RG_F_Site_xtab<- xtabs(~Site+lcat5, data=Females)
chisq.test(RG_F_Site_xtab)

RG_F_Month_xtab<- xtabs(~Month+lcat5, data=Females)
chisq.test(RG_F_Month_xtab)
```


## 3. Sex ratio calculation, visualization and statistical analysis

### Between months
```{r, echo=FALSE}
library(ratios)
options(digits = 3)

MF<- AllRG %>% filter(Sex != "I")

Sex_ratio<- xtabs(~  Month + Sex, data=MF)
Sex_ratio<- as.data.frame.matrix(Sex_ratio)



Sex_ratio$ratio<- case_when(
    is.na(Sex_ratio$M) & is.na(Sex_ratio$F) ~ 1,
    is.na(Sex_ratio$M) & Sex_ratio$F >= 0 ~ 1,
    Sex_ratio$M >= 0 & is.na(Sex_ratio$F) ~ 0,
    TRUE ~ Sex_ratio$M / Sex_ratio$F) 
colnames(Sex_ratio) <- c("Female", "Male", "Ratio")

Month<- rownames(Sex_ratio)
rownames(Sex_ratio)<- NULL
(Sex_ratio<- cbind(Month, Sex_ratio) %>% as_data_frame())

```


### Rose

```{r, echo=FALSE}

Rose_MF<- MF %>% filter(Site == "Rose")

Rose_MF_Sex_ratio<- xtabs(~  Month + Sex, data=Rose_MF)
Rose_MF_Sex_ratio<- as.data.frame.matrix(Rose_MF_Sex_ratio)

Rose_MF_Sex_ratio$ratio<- case_when(
    is.na(Rose_MF_Sex_ratio$M) & is.na(Rose_MF_Sex_ratio$F) ~ 1,
    is.na(Rose_MF_Sex_ratio$M) & Rose_MF_Sex_ratio$F >= 0 ~ 1,
    Rose_MF_Sex_ratio$M >= 0 & is.na(Rose_MF_Sex_ratio$F) ~ 0,
    TRUE ~ Rose_MF_Sex_ratio$M / Rose_MF_Sex_ratio$F) 
colnames(Rose_MF_Sex_ratio) <- c("Female.Rose", "Male.Rose", "Ratio.Rose")

Month<- rownames(Rose_MF_Sex_ratio)
rownames(Rose_MF_Sex_ratio)<- NULL
(Rose_MF_Sex_ratio<- cbind(Month, Rose_MF_Sex_ratio) %>% as_data_frame())

```

### Cobb

```{r, echo=FALSE}

Cobb_MF<- MF %>% filter(Site == "Cobb")
Cobb_MF_Sex_ratio<- xtabs(~  Month + Sex, data=Cobb_MF)
Cobb_MF_Sex_ratio<- as.data.frame.matrix(Cobb_MF_Sex_ratio)

Cobb_MF_Sex_ratio$ratio<- case_when(
    is.na(Cobb_MF_Sex_ratio$M) & is.na(Cobb_MF_Sex_ratio$F) ~ 1,
    is.na(Cobb_MF_Sex_ratio$M) & Cobb_MF_Sex_ratio$F >= 0 ~ 1,
    Cobb_MF_Sex_ratio$M >= 0 & is.na(Cobb_MF_Sex_ratio$F) ~ 0,
    TRUE ~ Cobb_MF_Sex_ratio$M / Cobb_MF_Sex_ratio$F) 
colnames(Cobb_MF_Sex_ratio) <- c("Female.Cobb", "Male.Cobb", "Ratio.Cobb")

Month<- rownames(Cobb_MF_Sex_ratio)
rownames(Cobb_MF_Sex_ratio)<- NULL
(Cobb_MF_Sex_ratio<- cbind(Month, Cobb_MF_Sex_ratio) %>% as_data_frame())
```

```{r}
(All_Sex_ratio<- cbind(Sex_ratio, Rose_MF_Sex_ratio, Cobb_MF_Sex_ratio))
All_Sex_ratio <- All_Sex_ratio[, !duplicated(colnames(All_Sex_ratio))]

```

```{r}
gt(All_Sex_ratio)
```



### Chi Square test 
  to detect differences in sex ratio between months
```{r, echo=FALSE}
library(RVAideMemoire)

chisq.test(Sex_ratio$Ratio)

(Sex_Chi<- chisq.theo.multcomp(Sex_ratio$Ratio, p.method = "bonferroni"))

(Sex_Chi<- extract(Sex_Chi))

(Sex_ratio<- cbind(Month, Sex_ratio, Sex_Chi))

Sex_ratio<- select(Sex_ratio, c("Month", "Female", "Male", "Ratio", "p.value.Chi", "p.value.Pr(>Chi)"))

```

### Summary table of sex ratio across sampling months

```{r, echo=FALSE}

(Sex_ratio_table<- gt(Sex_ratio) %>% cols_label(
    Month="Month",
    Male ="Male",
    Female="Female", 
    Ratio ="M:F Ratio",
    "p.value.Chi"= md("**X<sup>2</sup>**"),
     "p.value.Pr(>Chi)" = md("**P-value**")) %>% tab_options(column_labels.font.weight = "bold", row_group.font.weight = "bold") )
 
Sex_ratio_table %>% gtsave(filename="Sex_ratio_table.png", path="G:/Dissertation/Chapter 2- pop structure/RG_POP_STRUCTURE/Figures")



```


